A Structural and Functional Bioinformatics Study of QTY-designed Retinylidene Proteins

### 2025-05-09 Paper/Source/manuscript.tex
Complete manuscript text. Add graphs.

This study investigates the QTY design of 9 human and 3 microbial retinylidene proteins ... [TO BE WRITTEN]

@@

\textbf{Animal opsins} are a type of class A GPCR (G-protein coupled receptor), which are characterized by 7 transmembrane domains, the NPxxY motif, and activation of G-proteins through the outward movement of TM6. \citep{Zhou_2019}. Almost all animal opsins include a lysine residue that forms a Schiff base link with the retinal \citep{Guhmann_2022}. When retinal absorbs a photon, it isomerizes, usually changing from 11-cis to all-trans. The subsequent conformational changes of the protein are well studied using bovine rhodopsin, which changes from dark state to BATHO state, then LUMI, META I, and META II \citep{Okada_2001}. Proton transfer plays a major role in this process \citep{Mahalingam_2008}. During this process, TM6 is characteristically moved outward, activating the G-protein.

@@

\textbf{OPN2} (Opsin 2), also known as rhodopsin, is expressed in retinal rod photoreceptors and is responsible for vision at low light intensity \citep{Hubbard_1958}. Certain variants lead to autosomal recessive or autosomal dominant retinitis pigmentosa and congenital stationary night blindness \citep{Fanelli_2021}. OPN2 is a representative animal opsin, one of the earliest studied. In fact, bovine OPN2 is the first opsin to be sequenced \citep{Nathans_1984} as well as the first GPCR whose crystal structure was resolved experimentally \citep{Palczeski_2000}. Many studies on the functional mechanisms of animal opsins also focus on OPN2. Consequently, we also chose OPN2 to conduct a functional analysis in this study, in order to further explore the effectiveness of the QTY design of retinylidene proteins.

\textbf{OPN3} (Opsin 3), also known as encephalopsin or panopsin, is activated by blue and ultraviolet A light. It is expressed in the brain \citep{Blackshaw_1999}. It is also expressed in melanocytes and keratinocytes in the skin and regulates functions such as melanogenesis, cell differentiation, and glucose uptake. It is also expressed in the liver, pancreas, kidney, lung, heart, and skeletal muscles. When it is expressed in neurons, neurotransmitter release is inhibit by light, which renders OPN3 a useful optogenetic tool \citep{Copits_2021}.

\textbf{OPN4} (Opsin 4), also known as melanopsin, is expressed in ipRGC (intrinsically photosensitive Retinal Ganglion Cells) in the ganglion cell layer in the retina \citep{Provencio_1998}. It is essential for non-image-forming responses to light, including for pupillary reflex, optokinetic visual tracking response, and photoentrainment and regulation of circadian rhythm. Certain variants of OPN4 lead to seasonal affective disorder and other circadian rhythm disorders.  \citep{Berson_2002}. Image formation via OPN4-containing ipRGCs is also a potential pathway of treatment for various eye diseases such as retinitis pigmentosa and diabetic retinopathy.

@@

\textbf{RRH} (RPE-derived Rhodopsin Homolog), also known as peropsin, is localized in the microvilli of RPE cells that surround photoreceptor outer segments \citep{Sun_1997}. It is another protein that preferentially binds to all-trans-retinal \citep{Cook_2017}. Although not much information is known about RRH, it can reasonably be inferred that it photoisomerizes all-trans-retinal to 11-cis state, and may play a role in the upkeeping of photoreceptor functions.

\textbf{Bacterial opsins} are transmembrane ion pumps or channels \citep{Findlay_1986, Zhang_2011}. They are also 7TM, though this is due to convergent evolution rather than homology \citep{Yee_2013}. The isomerization of retinal is usually from all-trans to 13-cis, different from that in animal opsins \citep{Findlay_1986}. In addition, bacterial opsins often form oligomers to carry out their functions \citep{Gmelin_2007}.

In this study, we selected 3 bacterial opsins: BACR (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P02945/entry}{P02945}), BACH (UniProt ID: \href{https://www.uniprot.org/uniprotkb/B0R2U4/entry}{B0R2U4}), ChR2 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q8RUT8/entry}{Q8RUT8}).

\textbf{BACR} (Bacteriorhodopsin) is a light-driven proton pump and is one of the first bacterial opsins discovered \citep{Oesterhelt_1971}. \textbf{BACH} (Halorhodopsin) is a light-driven chloride pump activated by yellow light \citep{Schobert_1982}. \textbf{ChR2} (Channelrhodopsin 2) is a light-activated sodium channel activated by blue light \citep{Nagel_2003}. BACH and ChR2 are among the first optogenetic tools \citep{Zhang_2007, Han_2007}. BACH is used for inhibition, while ChR2 is used for excitation.

@@

Instead of taking a computational approach, we applied the QTY (Glutamine, Threonine, Tyrosine) code to systematically engineer water-soluble analogs with reduced hydrophobicity in membrane proteins. There are structural similarities between hydrophobic and polar amino acids: leucine (L) \textit{vs} glutamine (Q); isoleucine (I) / valine (V) \textit{vs} threonine (T); and phenylalanine (F) \textit{vs} tyrosine (Y), as can be observed on high-resolution electron density maps \citep{Zhang_2018, Zhang_2022, Tegler_2020}. This justifies the replacement of hydrophobic amino acids with polar ones. We initially applied the QTY code to design several detergent-free chemokines receptors, all of which retained structural thermal stability and native ligand-binding activities and enzymatic activities despite substantial changes to the transmembrane domain \citep{Zhang_2018, Tegler_2020}. We have also applied the QTY code to design water-soluble GPCRs, including chemokine receptors \citep{Zhang_2018, Qing_2019, Tegler_2020, Skuhersky_2021}, cytokine receptors \citep{Hao_2020}, and olfactory receptors \citep{Skuhersky_2021, Johnsson_2025}. Besides GPCRs, we have also used the QTY code to design water-soluble glucose transporters \citep{Smorodina_2022}, ABC transporters \citep{Pan_2024}, monoamine neurotransmitter transporters \citep{Karagol_Karagol_Alper_2024}, glutamate transporters \citep{Karagol_Karagol_Smorodina_2024, Karagol_2024}, mitochondrial megacomplex \citep{Chen_2025}, antibodies \citep{Li_Wang_2024}, potassium channels \citep{Smorodina_2024}, receptor kinases \citep{Li_Tang_2024}, transmembrane enzymes \citep{Chen_Pan_2025}, as well as bacterial enzymes with beta barrel structures \citep{Sajeev-Sheeja_2023, Sajeev-Sheeja_2024}. The QTY-designed water-soluble CXCR4 chemokine receptor has been successfully in biomimetic sensors \citep{Qing_2023}. We have also applied the rQTY (reverse QTY) code to design nanoparticles \citep{Meng_2023}.

AlphaFold2 was released by Google DeepMind in July 2021 \citep{Jumper_2021}. It greatly facilitated the study of QTY-designed protein variants. Subsequently, AlphaFold3 was released in May 2024, featuring improved architecture and improved efficiency \citep{Abramson_2024}. Furthermore, AlphaFold3 enabled the accurate prediction of complexes of multiple proteins, as well as complexes with nucleic acids and certain small molecules. QTY studies have made use of these new features \citep{Chen_2025, Johnsson_2025}.

GROMACS is a molecular dynamics (MD) simulation program released in 1995 \citep{Berendsen_1995}. Its 5.0 version was released in 2015 \citep{Abraham_2015}. GROMACS enables an efficient and realistic simulation of biomolecular systems, and has been used to investigate the structural and functional properties of QTY-designed protein variants in various studies \citep{Karagol_2024, Li_Tang_2024, Smorodina_2024, Li_Wang_2024, Johnsson_2025}.

@@

The native protein sequences for OPN1MW, OPN1LW, OPN1SW, OPN2, OPN3, OPN4, OPN5, RGR, RRH, BACR, BACH, and ChR2 were obtained from UniProt (\url{https://www.uniprot.org}). The sequence for ChR2 was truncated according to the experimentally determined structure in RCSB PDB (PDB ID: \href{https://www.rcsb.org/structure/8ZAN}{8ZAN}). The QTY designs were performed through the Protein Solubilizing Server (\url{https://pss.sjtu.edu.cn/}) \citep{Tao_2022}. For ChR2, the secondary structure was provided to the server in SS3 format according to the RCSB PDB structure.

@@

Regarding native OPN2, the membrane protein system was constructed using the web-based membrane builder CHARMM-GUI and was downloaded in GROMACS format \citep{Jo_2008, Wu_2014, Lee_2016}. The protein was centered in a rectangular box. The generated membrane models consisted of 40\% POPC, 40\% POPE, 10\% POPS, and 10\% cholesterol, which simulate a rod photoreceptor disk membrane \citep{Albert_2005}. The system was solvated in TIP3P water with 150 mM NaCl. 11-cis-retinal was manually added to the protein files. LINCS constraints were used for constraints, and Verlet integrator was used. Electrostatics was handled with Particle Mesh Ewald (PME), with both Coulomb and van der Waals interaction cutoffs set at 1.2 nm. The energy of the system was minimized using the steepest descent until the maximum forces converged below 1000 kJ/mol/nm. The standard six-step CHARMM-GUI NP$\gamma$T equilibration protocol \citep{Jo_2008} was used, with two 125-ps NVT equilibration simulations, and one 125-ps and three 500-ps NP$\gamma$T equilibration simulations. Temperature and pressure were maintained at 303.15 K and 1.0 bar, respectively, using the V-rescale thermostat and C-rescale barostat with surface-tension coupling. Following equilibration, a 10-ns production MD simulation was run. The retinal molecule was then manually switched to the all-trans state by rotating C13 to C15 and bonded hydrogen and methyl groups by 180 degrees around the C11-C12 bond. Energy minimization and equilibration were repeated exactly as before, and then a 200-ns production MD simulation was run.

Regarding QTY-designed OPN2, the system was constructed directly using GROMACS. The protein was centered in a rectangular box with dimensions equal to those of native OPN2. The equilibration involved a sequence of simulations: one 250-ps NVT equilibration, followed by one 125-ps and one 1500-ps NPT equilibrations, totaling an equilibration duration that matches that of the natural OPN2. All other configurations and parameters are identical to those of the original OPN2.

[ANALYSIS METHODS OF MD SIMULATION RESULTS, TO BE ADDED.]

@@

The protein sequences of 9 human opsins and 3 microbial opsins are aligned with their QTY analogs (Figure~\ref{fig:sequences}). The QTY substitution resulted in an overall variation of the sequence of 15.48\% to 30.04\% and a variation of transmembrane domain of 35.53\% to 50.24\%. Despite changes to the amino acid composition and sequence, the pI only experiences a slight change, between 0.00 and 0.16. This is because the amino acids Q, T, and Y have neutral charges and do not introduce additional charges to the system.  The MW increases slightly by a value between 0.04kDa and 0.60kDa. This is because the hydrophilic amino acids usually have nitrogen or oxygen atoms in place of carbon atoms in hydrophobic acids and thus may have higher mass: L (131.17Da) \textit{vs} Q (146.14Da); I (131.17Da) and V (117.15Da) \textit{vs} T (119.12Da); F (165.19Da) \textit{vs} Y (181.19Da).

@@

Although the chromophore, retinal, is bound to retinylidene proteins in an internal pocket, it is only incorporated into the protein after protein folding is complete. In other words, protein folding does not depend on the chromophore. Thus, it is justified to predict the ligand-free forms of the proteins. We used AlphaFold3 to predict the structure of the 9 native human opsins and their water-soluble QTY analogs. The structures superpose very well (Figure~\ref{fig:humansup} \textbf{a)} to \textbf{i)}). For OPN2, which has an available X-ray diffraction structure, we also superposed the experimentally determined structure with the AlphaFold3-predicted structures. Overall, the root mean square distances (RMSD) are small, from 0.307\AA to 0.611\AA, with only one exception of OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$ (RMSD = 0.999). This shows that the water-soluble QTY analogs are quite similar to native proteins in terms of structure.

@@

We investigate the surface hydrophobicity of water-soluble QTY-analogs in comparison with the native proteins. After applying the QTY code to respectively replace the hydrophobic amino acids L, I/V, and F with hydrophilic amino acids glutamine (Q), threonine (T), and tyrosine (Y), the hydrophobic surface areas are significantly reduced. More importantly, since the electron density maps of the corresponding amino acids are similar, the alpha helices in QTY analogs have a similar surface contour to the helices in the native proteins. The QTY analogs have successfully retained its structural integrity and stability.

@@

However, AlphaFold3 still has limitations that have been encountered in this study. Although AlphaFold3 allows the structural prediction of complexes that include small-molecule ligands, it only supports a small range of ligands, which does not include retinal. We were therefore unable to predict the structure of chromophore-bound states of retinylidene proteins. Consequently, we chose another approach, namely molecular dynamics simulation, in order to investigate the behavior of QTY opsin analogs when they are bound to the chromophore.

\subsection*{Simulation of 11-cis to all-trans isomerization of retinal in native OPN2 and its QTY variant}

Describe, explain, discuss MD simulation results. // NOTE: there were no changes to the NPxxY motif or retinal-binding Lys; "The ionic environment of the RSB, defined by the residues of the binding pocket, dictates the spectral and kinetic characteristics of each individual protein." \citep{Fenno_2011}, so QTY variant may have different absorption peak.

MD Results

* different states - binding pocket residues + ligand orientation

* bond length, angle, dihedral, non-bonded interaction lengths graph

* by-helix rmsd

* by-residue IE

* hbond

* overall rmsd, gyrate, rmsf, IE

@@

Future scopes and potential applications.

@@

Conclusion

@@

>>>>

Retinylidene proteins are a large family of retinal-binding light-sensitive proteins, found in many organisms ranging from microbes to human. Microbial opsins have been utilized in optogenetics, while animal opsins are essential for vision and other non-visual light-dependent metabolic functions. However, all retinylidene proteins are membrane proteins with hydrophobic transmembrane domains, which makes it challenging to study them. In this structural and functional bioinformatics study, we used the QTY (Glutamine, Threonine, Tyrosine) code to design water-soluble QTY analogs of retinylidene proteins, including 9 human opsins and 3 microbial opsins. We provide the superpositions of the AlphaFold3-predicted hydrophobic native proteins and their water soluble QTY variants, and experimentally-determined structures when available. We also provide a comparison of the surface hydrophobicity of the variants. Furthermore, we run molecular dynamics (MD) simulations of native and QTY-designed OPN2 and analyze their response to the 11-cis to all-trans isomerization of the retinal chromophore. The findings of this study indicate that the QTY code can be used as a robust tool to design water-solubility retinylidene proteins. These have potential applications in protein studies, therapeutic treatments, and bioengineering.

@@

\textbf{Animal opsins} are a type of class A GPCR (G-protein coupled receptor), which are characterized by 7 transmembrane domains, the NPxxY motif, and activation of G-proteins through the outward movement of TM6. \citep{Zhou_2019}. Almost all animal opsins include a lysine residue that forms a Schiff base link with the retinal \citep{Guhmann_2022}. When retinal absorbs a photon, it isomerizes, usually changing from 11-cis to all-trans. The subsequent conformational changes of the protein are well studied using bovine rhodopsin, which changes from dark state to BATHO state, then LUMI, META I, and META II \citep{Okada_2001}. Proton transfer plays an important role in this process \citep{Mahalingam_2008}. As the protein conformation changes, TM6 characteristically moves outward, activating the G-protein.

@@

\textbf{OPN2} (Opsin 2), also known as rhodopsin, is expressed in retinal rod photoreceptors and is responsible for vision at low light intensity \citep{Hubbard_1958}. Certain variants lead to autosomal recessive or autosomal dominant retinitis pigmentosa and congenital stationary night blindness \citep{Fanelli_2021}. OPN2 is a representative animal opsin, one of the first studied. In fact, bovine OPN2 is the first opsin to be sequenced \citep{Nathans_1984}, as well as the first GPCR whose crystal structure was resolved experimentally \citep{Palczeski_2000}. Many studies on the functional mechanisms of animal opsins also focus on OPN2. Consequently, we choose OPN2 to conduct a functional analysis in this study, in order to further explore the effectiveness of the QTY code in redesigning retinylidene proteins.

\textbf{OPN3} (Opsin 3), also known as encephalopsin or panopsin, is activated by blue and ultraviolet A light. It was discovered in the brain \citep{Blackshaw_1999}. It is also expressed in melanocytes and keratinocytes in the skin and regulates functions such as melanogenesis, cell differentiation, and glucose uptake. Its expression is also found in the liver, pancreas, kidney, lung, heart, and skeletal muscles. When expressed in neurons, the release of neurotransmitters is inhibited by light, making OPN3 a useful inhibitory optogenetic tool \citep{Copits_2021}.

\textbf{OPN4} (Opsin 4), also known as melanopsin, is expressed in ipRGC (intrinsically photosensitive Retinal Ganglion Cells) in the ganglion cell layer in the retina \citep{Provencio_1998}. It is essential for non-image-forming responses to light, including the pupillary reflex, optokinetic visual tracking response, and photoentrainment and regulation of circadian rhythm. Certain variants of OPN4 lead to seasonal affective disorder and other circadian rhythm disorders \citep{Berson_2002}. Rendering OPN4-containing ipRGCs capable of image formation is also a potential pathway for the treatment of various eye diseases such as retinitis pigmentosa and diabetic retinopathy.

@@

\textbf{RRH} (RPE-derived Rhodopsin Homolog), also known as peropsin, is localized in the microvilli of RPE cells that surround photoreceptor outer segments \citep{Sun_1997}. It is another protein that preferentially binds to all-trans-retinal \citep{Cook_2017}. Although not much information is known about RRH, it can reasonably be inferred that it photoisomerizes all-trans-retinal to 11-cis-retinal and may play a role in the upkeep of photoreceptor functions.

\textbf{Microbial opsins} are transmembrane ion pumps or channels \citep{Findlay_1986}. They are also 7TM, though this is due to convergent evolution rather than homology \citep{Yee_2013}. The chromophore, retinal, usually isomerizes from all-trans to 13-cis, different from the case of animal opsins \citep{Findlay_1986}. In addition, microbial opsins often form oligomers to carry out their functions \citep{Gmelin_2007}.

In this study, we selected 3 microbial opsins: BACR (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P02945/entry}{P02945}), BACH (UniProt ID: \href{https://www.uniprot.org/uniprotkb/B0R2U4/entry}{B0R2U4}), ChR2 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q8RUT8/entry}{Q8RUT8}).

\textbf{BACR} (Bacteriorhodopsin) is a light-driven proton pump. It is one of the first microbial opsins discovered \citep{Oesterhelt_1971}. \textbf{BACH} (Halorhodopsin) is a light-driven chloride pump activated by yellow light \citep{Schobert_1982}. \textbf{ChR2} (Channelrhodopsin 2) is a light-activated sodium channel activated by blue light \citep{Nagel_2003}. BACH and ChR2 are among the first optogenetic tools \citep{Zhang_2007, Han_2007}. BACH is used for inhibition, while ChR2 is used for excitation.

@@

Instead of taking a computational approach, we applied the QTY (Glutamine, Threonine, Tyrosine) code to systematically engineer water-soluble analogs with reduced hydrophobicity in membrane proteins. There are structural similarities between hydrophobic and polar amino acids: leucine (L) and glutamine (Q); isoleucine (I) / valine (V) and threonine (T); and phenylalanine (F) and tyrosine (Y), as can be observed on high-resolution electron density maps \citep{Zhang_2018, Zhang_2022, Tegler_2020}. This justifies the replacement of hydrophobic amino acids with polar ones. We initially applied the QTY code to design several detergent-free chemokines receptors, all of which retained structural thermal stability and native ligand-binding and enzymatic activities despite substantial changes to the transmembrane domain \citep{Zhang_2018, Tegler_2020}. We have also applied the QTY code to design water-soluble GPCRs, including chemokine receptors \citep{Zhang_2018, Qing_2019, Tegler_2020, Skuhersky_2021}, cytokine receptors \citep{Hao_2020}, and olfactory receptors \citep{Skuhersky_2021, Johnsson_2025}. In addition to GPCR, we have also used the QTY code to design water-soluble glucose transporters \citep{Smorodina_2022}, ABC transporters \citep{Pan_2024}, monoamine neurotransmitter transporters \citep{Karagol_Karagol_Alper_2024}, glutamate transporters \citep{Karagol_Karagol_Smorodina_2024, Karagol_2024}, mitochondrial megacomplex \citep{Chen_2025}, antibodies \citep{Li_Wang_2024}, potassium channels \citep{Smorodina_2024}, receptor kinases \citep{Li_Tang_2024}, transmembrane enzymes \citep{Chen_Pan_2025}, as well as bacterial enzymes with beta barrel structures \citep{Sajeev-Sheeja_2023, Sajeev-Sheeja_2024}. The QTY-designed water-soluble CXCR4 chemokine receptor has been successfully used in biomimetic sensors \citep{Qing_2023}. We have also applied the rQTY (reverse QTY) code to design self-assembled nanoparticles \citep{Meng_2023}.

AlphaFold2 was released by Google DeepMind in July 2021 \citep{Jumper_2021}. It has greatly facilitated the study of QTY-designed protein variants. Subsequently, AlphaFold3 was released in May 2024, featuring improved architecture and improved efficiency \citep{Abramson_2024}. Furthermore, AlphaFold3 enabled the accurate prediction of complexes of multiple proteins, as well as complexes with modified residues, nucleic acids, ions, certain ligands. QTY studies have made use of these new features \citep{Chen_2025, Johnsson_2025}.

GROMACS is a molecular dynamics (MD) simulation program released in 1995 \citep{Berendsen_1995}. Its 5.0 version was released in 2015 \citep{Abraham_2015}. GROMACS enables an efficient and realistic simulation of biomolecular systems and has been used to investigate the structural and functional properties of QTY-designed protein variants in various studies \citep{Karagol_2024, Li_Tang_2024, Smorodina_2024, Li_Wang_2024, Johnsson_2025}.

@@

The native protein sequences for OPN1MW, OPN1LW, OPN1SW, OPN2, OPN3, OPN4, OPN5, RGR, RRH, BACR, BACH, and ChR2 were obtained from UniProt (\url{https://www.uniprot.org}). The sequence for ChR2 was truncated according to the experimentally determined structure in the RCSB PDB (PDB ID: \href{https://www.rcsb.org/structure/8ZAN}{8ZAN}). The QTY designs were performed through the Protein Solubilizing Server (\url{https://pss.sjtu.edu.cn/}) \citep{Tao_2022}. For ChR2, the secondary structure was provided to the server in SS3 format according to the RCSB PDB structure.

@@

Regarding native OPN2, the membrane protein system was constructed using the web-based membrane builder CHARMM-GUI and was downloaded in GROMACS format \citep{Jo_2008, Wu_2014, Lee_2016}. The protein was centered in a rectangular box. The generated membrane models consisted of 40\% POPC, 40\% POPE, 10\% POPS, and 10\% cholesterol, which simulate a rod photoreceptor disk membrane \citep{Albert_2005}. The system was solvated in TIP3P water with 150 mM NaCl. 11-cis-retinal was manually added to the protein files. LINCS constraints were used for constraints, and Verlet integrator was used. Electrostatics was handled with Particle Mesh Ewald (PME), with both Coulomb and van der Waals interaction cutoffs set at 1.2 nm. The energy of the system was minimized using the steepest descent until the maximum forces converged below 1000 kJ/mol/nm. The standard six-step CHARMM-GUI NP$\gamma$T equilibration protocol \citep{Jo_2008} was used, with two 125-ps NVT equilibration simulations, and one 125-ps and three 500-ps NP$\gamma$T equilibration simulations. Temperature and pressure were maintained at 303.15 K and 1.0 bar, respectively, using the V-rescale thermostat and C-rescale barostat with surface-tension coupling. Following equilibration, a 10-ns production MD simulation was run. The retinal molecule was then manually switched to the all-trans state by rotating C13 to C15 and bonded hydrogen and methyl groups by 180 degrees around the C11-C12 bond. Energy minimization and equilibration were repeated exactly as before, and then a 120-ns production MD simulation was run.

Regarding QTY-designed OPN2, we found that the helix 9 of the QTY analog was flung up and adhered to the originally transmembrane domain, which disrupted the normal conformation. Therefore, this helix was truncated. The system was constructed directly using GROMACS. The protein was centered in a rectangular box with dimensions equal to those of native OPN2. The equilibration involved a sequence of simulations: one 250-ps NVT equilibration, followed by one 125-ps and one 1500-ps NPT equilibrations, totaling an equilibration duration that matches that of the natural OPN2. All other configurations and parameters are identical to those of the original OPN2.

After the simulations, we extracted frames from the trajectory to inspect the structures before and after retinal isomerization. We also calculated the RMSD (``gmx rms'' command), number of hydrogen bonds (``gmx hbond'' command) and water molecules (``gmx select'' command) inside the binding pocket, interaction energy (including short-range Coulombic interaction energy and short-range Lennard-Jones energy) (``gmx energy'' command), the RMSF (``gmx rmsf'' command), and radius of gyration (``gmx gyrate'' command). The detailed commands are publicly available in the online database for this study.

@@

The protein sequences of 9 human opsins and 3 microbial opsins are aligned with their QTY analogs (Figure~\ref{fig:sequences}). The QTY substitution resulted in an overall variation of the sequence of 15.48\% to 30.04\% and a variation of transmembrane domain of 35.53\% to 50.24\%. Despite changes in amino acid composition and sequence, the pI only experiences a slight change, between 0.00 and 0.16. This is because the amino acids Q, T, and Y have neutral charges and do not introduce additional charges to the system.  The MW increases slightly by a value between 0.04kDa and 0.60kDa. This is because the hydrophilic amino acids usually have nitrogen or oxygen atoms in place of carbon atoms in hydrophobic acids and thus may have higher mass: L (131.17Da) \textit{vs} Q (146.14Da); I (131.17Da) and V (117.15Da) \textit{vs} T (119.12Da); F (165.19Da) \textit{vs} Y (181.19Da).

@@

Although the chromophore, retinal, is bound to retinylidene proteins in an internal pocket, it is only incorporated into the protein after protein folding is complete. In other words, protein folding does not depend on the chromophore. Thus, it is justified to predict the ligand-free forms of the proteins. We used AlphaFold3 to predict the structure of the 9 native human opsins and their water-soluble QTY analogs. The structures superpose very well (Figure~\ref{fig:humansup} \textbf{a)} to \textbf{i)}). For OPN2, which has an available X-ray diffraction structure, we also superposed the experimentally determined structure with the AlphaFold3-predicted structures. Overall, the root mean square distances (RMSD) are small, from 0.307\AA to 0.611\AA, with only one exception of OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$ (RMSD = 0.999\AA). This shows that the water-soluble QTY analogs are quite similar to native proteins in terms of structure.

@@

We investigate the surface hydrophobicity of water-soluble QTY-analogs in comparison with the native proteins. After applying the QTY code to respectively replace the hydrophobic amino acids L, I/V, and F with hydrophilic amino acids glutamine (Q), threonine (T), and tyrosine (Y), the hydrophobic surface areas are significantly reduced. More importantly, since the electron density maps of the corresponding amino acids are similar, the alpha helices in QTY analogs have a similar contour to those in the native proteins. The QTY analogs have successfully retained its structural integrity and stability.

@@

However, AlphaFold3 still has limitations that have been encountered in this study. Although AlphaFold3 allows the structural prediction of complexes that include small-molecule ligands, it only supports a small range of ligands, which does not include retinal. We were therefore unable to predict the structure of chromophore-bound states of retinylidene proteins with AlphaFold3. Consequently, we chose another approach, namely molecular dynamics simulation, in order to investigate the behavior of QTY opsin analogs when they are bound to the chromophore.

\subsection*{Simulation of 11-cis to all-trans isomerization of retinal in native OPN2 and its QTY analog}

We observe that QTY substitutions do not introduce changes to the essential NPxxY motif or the retinal-binding lysine of human retinylidene proteins. Therefore, we propose that QTY analogs of human opsins may have conserved functions and carry out a molecular dynamics (MD) simulation. We record and analyze the behavior of native OPN2 and its QTY analog before and after the 11-cis to all-trans isomerization.

The MD simulation provides evidence of conformational change in both native and QTY analogs (Figure~\ref{fig:conformation}). The native protein experiences a slower change from its original conformation, while the QTY analog shows more fluctuations and sudden changes. Under further inspection, this phenomenon may be due to the entrance of water molecules into the QTY analog. In addition, retinal molecule has a greater degree of change in conformation in the native protein than in the QTY analog. Nevertheless, the retinal in the QTY analog is still able to cause changes to surrounding residues, sometimes doing so through contact with water molecules.

Zooming in on the retinal-binding pocket, we observe that retinal is held within its binding pocket by several forces: the counterions of the retinal Schiff base (RSB) (113E and 181E), the hydrophobic interactions with neighboring residues (for example, 265W, 268Y, and the 187--189 beta strand), and, sometimes, steric collisions (Figure~\ref{fig:binding}). Since retinal is a ligand that binds in an internal pocket, we were unable to calculate the Gibbs free energy of its binding. Nevertheless, the distribution of interaction energies are found to be roughly the same in both native OPN2 and the water-soluble QTY analog, which is a strong suggestion of functional conservation. Finally, we observe very different hydrogen bonding and water molecule patterns inside the binding pocket, which is mainly due to several F to Y substitutions in that region. This may cause retinal to interact with adjacent residues in a subtly different way, and may lead to a different absorption peak since the environment of the RSB, i.e., the binding pocket, determines the spectral characteristic of opsins \citep{Fenno_2011}.

Overall, we observe that QTY-designed OPN2 has relatively similar behavior to the native variant despite certain small differences caused by decreased hydrophobicity. The function of OPN2$^{\textrm{QTY}}$ awaits to be verified by experiments.

@@

The findings of this study indicate that the QTY code could be used as a robust tool to design water-soluble retinylidene proteins.

This could, in the first place, facilitate the study of these proteins. The water-solubility of QTY analogs makes it easier to purify and investigate proteins, especially those such as RGR and RRH, which have not been studied much but may have considerable clinical relevance. The water-solubility also reduces the difficulty in recording the different phases of the photocycle, in comparison with native membrane opsins. Another application may be the rapid design of new optogenetic tools.

Furthermore, water-soluble QTY opsins may be useful clinically, since they could be delivered and/or expressed in the eye without forming aggregates and precipitating. This may be a potential pathway in restoring basic vision for those who have lost it.

Last but not least, water-soluble opsins have the potential to be harvested in mass and may be used to design new biomimetic light-sensing systems, which may have applications in bioengineering.

@@

In this study, we selected 12 retinylidene proteins, including 9 human opsins (OPN1MW, OPN1LW, OPN1SW, OPN2, OPN3, OPN4, OPN5, RGR, and RRH) and 3 microbial opsins (BACR, BACH, and ChR2), that have clinical implications and various applications. We applied the QTY code to convert the hydrophobic transmembrane alpha-helices into hydrophilic ones and thus create water-soluble QTY analogs of the proteins. Then, we used AlphaFold3 to predict the structures of the native proteins as well as the QTY analogs, superposed them, and found that the structure of the proteins are well conserved despite substantial residue substitutions. We also inspected the surface hydrophobicity of the proteins and found a great decrease in hydrophobic patches in the transmembrane surface. Next, we chose a representative retinylidene protein, OPN2, and employed molecular dynamics simulations to investigate the behavior of native and QTY analog of OPN2. We found that the QTY analog is capable of conformational change and effective protein-ligand interaction. These findings reveal that QTY-designed OPN2, and retinylidene proteins in general, may have conserved structure and function. We believe that these water-soluble QTY variants may have potentials in research, therapeutic treatments, and bioengineering.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/function-conformation.jpg}
	\caption{\textbf{The conformational changes of native OPN2 and its QTY analog before and after 11-cis to all-trans isomerization of the chromophore, retinal. } \textbf{a)} and \textbf{b)}: 1ns running averages of the root mean square distances (RMSD) of the protein-retinal complex, transmembrane helix 6 (TM6), the retinal-binding pocket, and retinal. \textbf{c)}: superpositions between cis-state OPN2, trans-state OPN2, cis-state OPN2$^{\textrm{QTY}}$, and trans-state OPN2$^{\textrm{QTY}}$. Blue: cis-state protein, orange: 11-cis-retinal; yellow: trans-state protein, greenish cyan: all-trans retinal. }
	\label{fig:conformation}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{Figures/function-binding.jpg}
	\caption{\textbf{Changes in the retinal-binding pocket and protein-ligand interaction in native OPN2 and its QTY analog before and after 11-cis to all-trans isomerization of the chromophore. } \textbf{a)} and \textbf{b)}: close-ups of the binding pocket in cis-state. Blue: protein residues, orange: 11-cis-retinal; dashed lines: same as a) and b). \textbf{c)} and \textbf{d)}: close-ups of the binding pocket in trans-state. Yellow: protein residues, greenish cyan: all-trans-retinal; green dashed lines: ion bridge; yellow dashed lines: van der Waals and/or hydrophobic interactions. \textbf{e)} and \textbf{f)}: the interaction energies between the protein, the binding pocket, individual residues, and retinal. The interaction energy is the sum of the short-range Coulombic interaction energy and short-range Lennard-Jones energy. Please note that the interaction energy is a product of MD simulation and is not necessarily a meaningful physical quantity. \textbf{g)} and \textbf{h)}: the number of hydrogen bonds formed between residues in the binding pocket, and the number of water molecules within the binding pocket. }
	\label{fig:binding}
\end{figure}
